Price of Anarchy for Mean Field Games

### 2018-08-03 PoA.tex
minor edits

\subsubsection*{\textbf{Social Cost}}
We restrict ourselves to Markovian control strategies $\balpha=(\alpha_t)_{0\le t\le T}$ given by feedback functions in the form $\alpha_t=\phi(t,X_{t})$ and we let $\mathbb{A}$ denote the set of such controls. If the $N$ players use distributed Markovian control strategies of the form $\alpha^i_t=\phi(t,X^i_t)$, we define the cost (per player) to the system as the quantity $J^{(N)}_\phi$:
\begin{equation*}
J^{(N)}_\phi=\frac1N\sum_{i=1}^NJ^i(\balpha^1,\cdots,\balpha^N).
\end{equation*}
We shall compute this social cost in the limit $N\to\infty$ when all the players use the distributed control strategies given by the same feedback function $\phi$ identified by solving an optimization problem in the limit $N\to\infty$. We take the social cost to be the limit as $N\to\infty$ of $J^{(N)}_\phi$, namely
\begin{equation*}
\begin{split}
\lim_{N\to\infty}J^{(N)}_\phi
&=\lim_{N\to\infty}\frac1N\sum_{i=1}^NJ^i(\balpha^1,\cdots,\balpha^N)\\
&=\lim_{N\to\infty}\frac1N\sum_{i=1}^N\EE\bigg[\int_0^Tf(t,X^i_t,\mu^{N}_{t},\phi(t,X^i_t),\nu^{N}_{t})\,dt +g(X^i_T,\mu^{N}_{T})
\bigg],\\
&=\lim_{N\to\infty}\EE\bigg[\int_0^T<f(t,\,\cdot\,,\mu^{N}_{X_t},\phi(t,\,\cdot\,),\nu^{N}_{t}),\mu^N_{t}>\,dt +<g(\,\cdot\,,\mu^{N}_{X_T}),\mu^{N}_{T}>
\bigg],
\end{split}
\end{equation*}
if we use the notation $<\varphi,\rho>$ for the integral $\int \varphi(z)\rho(dz)$ of the function $\varphi$ with respect to the measure $\rho$. Now if we assume that in the limit $N\to\infty$ the empirical distributions $\mu^N_{t}$ converge toward a measure $\mu_t$, and thus $\nu^N_{t}=\frac1N\sum_{i=1}^N\delta_{\phi(t,X^i_t)}$ also converges toward a measure $\nu_t$, then the social cost of the feedback function $\phi$ becomes:
\begin{equation*}
SC(\phi)=\int_0^T<f(t,\,\cdot\,,\mu_t,\phi(t,\,\cdot\,),\nu_t),\mu_t>\,dt \;+\; <g(\,\cdot\,,\mu_T),\mu_T>
\end{equation*}
with the expectation, $\EE$, disappearing when the limiting flows $\bmu=(\mu_t)_{0\le t\le T}$ and $\bnu=(\nu_t)_{0\le t\le T}$ are deterministic.

@@

\subsection{\textbf{Linear Quadratic Extended Mean Field Games}}\label{sec:EMFG}
To solve the linear quadratic extended mean field game (LQEMFG), we begin by considering the reduced Hamiltonian for this problem:
\begin{equation*}
\begin{split}
    H(t,x,\bar{\mu},\alpha,\bar{\nu},y)=&\left[b_1(t)x+\bar{b}_1(t)\bar{\mu}+b_2(t) \alpha+\bar{b}_2(t)\bar{\nu}\right]y \\
    &+\frac{1}{2}\left[q(t)x^2+\bar{q}(t)(x-s(t)\bar{\mu})^2 +r(t)\alpha^2+\bar{r}(t)(\alpha-\bar{s}(t)\bar{\nu})^2\right],
\end{split}
\end{equation*}
and whenever the flows $\bmubar=(\bar{\mu}_t)_{0\leq t\leq T}$ and $\bnubar=(\bar{\nu}_t)_{0\leq t\leq T}$ are fixed, we consider for each control process $\balpha=(\alpha_t)_{0\leq t\leq T}$ the adjoint equation:
\begin{equation*}
    dY_t=- \partial_x H(t,X_t,\bar{\mu}_t,\alpha_t,\bar{\nu}_t,Y_t)dt+Z_tdW_t,
    \qquad
    Y_T=\partial_xg(X_T, \cL(X_T)).
\end{equation*}
According to the Pontryagin stochastic maximum principle, a sufficient condition for optimality is $\partial_{\alpha}H(t,X_t,\bar{\mu}_t,\hat{\alpha},\bar{\nu}_t,Y_t)=0$. Thus, we find the optimal control:
\begin{equation}
    \hat{\alpha}_t=\hat{\alpha}(t,X_t,\bar{\mu},\bar{\nu},Y_t)=\frac{\bar{r}(t)\bar{s}(t)\bar{\nu}-b_2(t)Y_t}{r(t)+\bar{r}(t)},
\label{eq:MFG_opt_alpha_Pontryagin}
\end{equation}
When solving the fixed point step, we identify $\bar{\nu}_t=\mathbb{E}(\hat{\alpha}_t)$. By taking the expectation, we find:
\begin{equation*}
    \bar{\nu}_t = \mathbb{E}(\hat{\alpha}_t)=c^{MFG}(t)\mathbb{E}(Y_t),
	\qquad 
	\text{with} 
	\quad
    c^{MFG}(t)=-\frac{b_2(t)}{r(t)+\bar{r}(t)(1-\bar{s}(t))}.
\end{equation*}
Thus, from equation (\ref{eq:MFG_opt_alpha_Pontryagin}) we have:
\begin{equation}
    \hat{\alpha}_t=a^{MFG}(t) Y_t+b^{MFG}(t)\mathbb{E}(Y_t),
\label{eq:MFG_opt_alpha_Y_expY}
\end{equation}
with:
\begin{equation*}
	a^{MFG}(t)=-\frac{b_2(t)}{r(t)+\bar{r}(t)},
	\qquad
	\text{and}
	\qquad
	b^{MFG}(t)=-\frac{\bar{r}(t)\bar{s}(t)b_2(t)}{(r(t)+\bar{r}(t))(r(t)+\bar{r}(t)(1-\bar{s}(t)))}.
\end{equation*}
Note that $c^{MFG}(t)=a^{MFG}(t)+b^{MFG}(t)$. The solution of the mean field game equilibrium problem is given by the solution to the FBSDE system:
\begin{equation}
\left\{
\begin{array}{lcl}
      \displaystyle  dX_t&=& \left[b_1(t)X_t+\bar{b}_1(t) \mathbb{E}X_t+a^{MFG}(t)b_2(t)Y_t+(b^{MFG}(t)b_2(t)+c^{MFG}(t)\bar{b}_2(t))\mathbb{E}Y_t\right]dt+\sigma dW_t \\[5pt]
        dY_t&=& -\left[(q(t)+\bar{q}(t))X_t-\bar{q}(t)s(t)\mathbb{E}X_t+b_1(t)Y_t \right]dt +Z_t dW_t,
\end{array}
\right.
\label{eq:FBSDE_EMFG}
\end{equation}
with initial condition $X_0=\xi$, a random variable with finite mean and variance, and terminal condition $Y_T=(q_T+\bar{q}_T)X_T-\bar{q}_Ts_T\mathbb{E}X_T$.

This is a linear FBSDE of McKean-Vlasov type, which can be solved explicitly under mild assumptions (or at least in the case of time-independent coefficients which we will consider later. See Appendix \ref{appendix_solving_fbsdes}). Let $\bar{\eta}_t^{MFG}$, $\eta_t^{MFG}$, $\bar{x}_t^{MFG}$, and $v^{MFG}_t$ denote the solutions for this problem as described in the appendix so that:
\begin{align*}
Y_t =\eta_t^{MFG}X_t+(\bar{\eta}_t^{MFG}-\eta_t^{MFG})\bar{x}_t^{MFG}, \quad &\mathbb{E}(Y_t) =\bar{\eta}_t^{MFG}\bar{x}_t^{MFG},\\
\mathbb{E}(X_t) =\bar{x}_t^{MFG}, \quad &Var(X_t)=v^{MFG},
\end{align*}
provide a solution to the LQEMFG problem. Then from the appendix, we have:
\begin{itemize}
	\item a scalar Riccati equation for $\bar{\eta}_t^{MFG}$:
	\begin{equation}
	\dot{\bar{\eta}}^{MFG}_t+ \left[ c^{MFG}(t)(b_2(t)+\bar{b}_2(t)) \right] \cdot (\bar{\eta}^{MFG}_t)^2+ \left(2b_1(t)+\bar{b}_1(t) \right) \cdot \bar{\eta}^{MFG}_t +q(t)+\bar{q}(t)(1-s(t)) =0
	\label{eq:eta_bar_MFG}
	\end{equation}
	with terminal condition $\bar{\eta}^{MFG}_T = q_T+\bar{q}_T(1-s_T)$, 
	\item a linear first order ODE for $\bar{x}_t^{MFG}$:
	\begin{equation}
	\dot{\bar{x}}^{MFG}_t=\left[ b_1(t)+\bar{b}_1(t)+c^{MFG}(t)(b_2(t)+\bar{b}_2(t)) \cdot \bar{\eta}^{MFG}_t \right] \cdot \bar{x}^{MFG}_t
	\label{eq:x_bar_MFG}
	\end{equation}
	with initial condition $\bar{x}^{MFG}_0=\mathbb{E}(\xi)$,
	\item a scalar Riccati equation for $\eta_t^{MFG}$:
	\begin{equation}
	\dot{\eta}^{MFG}_t+a^{MFG}(t)b_2(t)\cdot (\eta^{MFG}_t)^2 + 2b_1(t) \cdot \eta^{MFG}_t+q(t)+\bar{q}(t)=0
	\end{equation}
	with terminal condition $\eta^{MFG}_T = q_T+\bar{q}_T$,
\end{itemize}
where the dot is the standard ODE notation for a derivative. 
And thus, we obtain explicit solutions for $\bar{x}_t^{MFG}$ and $v_t^{MFG}$:
\begin{IEEEeqnarray}{rCl}
	\bar{x}^{MFG}_t &=& \mathbb{E}(\xi) e^{ \int_0^t\left( b_1(s)+\bar{b}_1(s)+ \left[ c^{MFG}(s)(b_2(s)+\bar{b}_2(s))\right] \cdot \bar{\eta}^{MFG}_s\right) ds} \label{eq:x_bar_MFG_explicit} \\ 
	v^{MFG}_t &=& Var(\xi)e^{\int_0^t 2\left[b_1(s)+a^{MFG}(s)b_2(s) \cdot \eta^{MFG}_s\right]ds}+\sigma^2 \int_0^t e^{2 \int_s^t \left[b_1(u)+a^{MFG}(u)b_2(u)\eta^{MFG}_u\right] du}ds.
	\label{eq:v_t}
\end{IEEEeqnarray}
Let $SC^{MFG}:=SC(\phi)$ in which $\phi=\phi(t,x)$ is the feedback function specified by this solution, namely, from equation (\ref{eq:MFG_opt_alpha_Y_expY}) we have:
\begin{equation*}
    \phi(t,x)=a^{MFG}(t)\eta_t^{MFG}x+\left[a^{MFG}(t)(\bar{\eta}_t^{MFG}-\eta_t^{MFG})+b^{MFG}(t)\bar{\eta}_t^{MFG} \right]\bar{x}_t^{MFG}.
\end{equation*}
Then we can compute the social cost as described in Section \ref{sub:MFG_formulation}:
\begin{equation*}
\begin{split}
    SC^{MFG}=&\frac{1}{2}\Big[(q_T+\bar{q}_T)v_T^{MFG}+(q_T+\bar{q}_T(1-s_T)^2)(\bar{x}_T^{MFG})^2\\
    &+\int_0^T \left[q(t)+\bar{q}(t)+(r(t)+\bar{r}(t))(a^{MFG}(t)\eta_t^{MFG})^2\right]v_t^{MFG}dt\\
    &+\int_0^T\left[q(t)+\bar{q}(t)(1-s(t))^2+(r(t)+\bar{r}(t)(1-\bar{s}(t))^2)(c^{MFG}(t)\bar{\eta}_t^{MFG})^2\right](\bar{x}_t^{MFG})^2dt\Big],
\end{split}
\end{equation*}
where we have used the fact that:
\begin{equation*}
    \mathbb{E}(\phi(t,X_t))=c^{MFG}(t)\cdot \bar{\eta}_t^{MFG}\bar{x}_t^{MFG}
	\qquad 
	\text{and}
	\qquad
    Var(\phi(t,X_t))=\left(a^{MFG}(t)\eta_t^{MFG} \right)^2 \cdot v^{MFG}_t.
\end{equation*}

\subsection{\textbf{Linear Quadratic Control of McKean-Vlasov Type Involving the Law of the Control}}\label{sec:EMKV}
To solve the linear quadratic optimal control problem of McKean-Vlasov type involving the law of the control (LQEMKV), we begin with the reduced Hamiltonian, which is the same as in the LQEMFG problem:
\begin{equation*}
\begin{split}
    H(t,x,\bar{\mu},\alpha,\bar{\nu},y)=&\left[b_1(t)x+\bar{b}_1(t)\bar{\mu}+b_2(t) \alpha+\bar{b}_2(t)\bar{\nu}\right]y\\
    &+\frac{1}{2}\left[q(t)x^2+\bar{q}(t)(x-s(t)\bar{\mu})^2 +r(t)\alpha^2+\bar{r}(t)(\alpha-\bar{s}(t)\bar{\nu})^2\right].
\end{split}
\end{equation*}
Since we require $\bar{\nu}_t$ to be equal to $\mathbb{E}(\alpha_t)$ throughout the optimization, it is not sufficient to minimize the Hamiltonian with respect to the $\alpha$ input alone in order to guarantee optimality. A sufficient condition for control problems of McKean-Vlasov type involving the law of the control is derived in \cite{carmona_acciaio}. Since we consider a Hamiltonian that depends on the means of $\bar{\mu}$ and $\bar{\nu}$ instead of the full distributions, the sufficient condition reduces to the following (see section 4 in \cite{carmona_acciaio}):
$\hat{\alpha}_t=\hat{\alpha}(t,X_t,\bar{\mu},\bar{\nu},Y_t)$ should satisfy:
\begin{equation*}
    \partial_{\alpha}H(t,X_t,\mathbb{E}(X_t),\hat{\alpha}_t,\mathbb{E}(\hat{\alpha}_t),Y_t)+\tilde{\mathbb{E}} \left[\partial_{\bar{\nu}}H(t,\tilde{X}_t,\mathbb{E}(X_t),\hat{\alpha}_t,\mathbb{E}(\hat{\alpha}_t),\tilde{Y}_t) \right]=0,
\end{equation*}
where the adjoint equation is given by:
\begin{equation*}
\left\{
\begin{array}{lcl}
    dY_t&=&-\left[\partial_x H(t,X_t,\bar{\mu}_t,\alpha_t,\bar{\nu}_t,Y_t)+\tilde{\mathbb{E}} \left[\partial_{\bar{\mu}}H(t,\tilde{X}_t,\bar{\mu}_t,\tilde{\alpha}_t,\bar{\nu}_t,\tilde{Y}_t) \right] \right]dt+Z_tdW_t \\[8pt]
    Y_T&=&\partial_xg(X_T, \cL(X_T))+\tilde{\mathbb{E}} \left[\partial_{\bar{\mu}}g(\tilde{X}_T,\cL(X_T))(X_T) \right],
\end{array}
\right.
\end{equation*}
where $(\boldsymbol{\tilde{X}},\boldsymbol{\tilde{Y}},\boldsymbol{\tilde{\alpha}})$ denotes an independent copy of $(\boldsymbol{X},\boldsymbol{Y},\boldsymbol{\alpha})$. In the present LQ case, the sufficient condition can be used to solve for:
\begin{equation}
    \hat{\alpha}_t=a^{MKV}(t)Y_t+b^{MKV}(t)\mathbb{E}(Y_t),
    \label{eq:alpha_hat_MKV}
\end{equation}
with:
\begin{equation*}
\begin{split}
    a^{MKV}(t)&=-\frac{b_2(t)}{r(t)+\bar{r}(t)}, \quad \text{and} \quad b^{MKV}(t)=-\frac{1}{r(t)+\bar{r}(t)}\left(\bar{b}_2(t)-\frac{\bar{r}(t)\bar{s}(t)(\bar{s}(t)-2)(b_2(t)+\bar{b}_2(t))}{r(t)+\bar{r}(t)(1-\bar{s}(t))^2} \right).
\end{split}
\end{equation*}
Then $\mathbb{E}(\hat{\alpha}_t)=c^{MKV}(t)\mathbb{E}(Y_t)$ with:
   $$ c^{MKV}(t)=a^{MKV}(t)+b^{MKV}(t)=-\frac{b_2(t)+\bar{b}_2(t)}{r(t)+\bar{r}(t)(1-\bar{s}(t))^2}.$$

@@

\subsection{\textbf{A Particular Example: Flocking}}
Mean field game models of flocking have been proposed in the literature by Nourian et al \cite{nourian2010synthesis}\cite{nourian2011mean}. Here we consider a slightly different formulation as described in Section 3.6.1 of the book \cite{Carmona_book}. A representative bird in the flock controls their velocity, $X_t$, through the drift:
\begin{equation*}
    b(t,x,\mu,\alpha,\nu)=\alpha.
\end{equation*}
Suppose the initial condition is $X_0=1$. They choose their control with two goals in mind: to minimize their effort put into their control, and to align their velocity with the average velocity of the group. Thus, they consider the cost functions given by:
\begin{equation*}
\begin{split}
    f(t,x,\mu,\alpha,\nu)&=\frac{1}{2}\left(|x-\bar{\mu}|^2 +\alpha^2 \right), \\
    g(x,\mu)&=0.
\end{split}
\end{equation*}

>>>>

\subsubsection*{\textbf{Social Cost}}
We restrict ourselves to Markovian control strategies $\balpha=(\alpha_t)_{0\le t\le T}$ given by feedback functions in the form $\alpha_t=\phi(t,X_{t})$ and we let $\mathbb{A}$ denote the set of such controls. If the $N$ players use distributed Markovian control strategies of the form $\alpha^i_t=\phi(t,X^i_t)$, we define the cost (per player) to the system as the quantity $J^{(N)}_\phi$:
\begin{equation*}
J^{(N)}_\phi=\frac1N\sum_{i=1}^NJ^i(\balpha^1,\cdots,\balpha^N).
\end{equation*}
We shall compute this social cost in the limit $N\to\infty$ when all the players use the distributed control strategies given by the same feedback function $\phi$ identified by solving an optimization problem in the limit $N\to\infty$. We take the social cost to be the limit as $N\to\infty$ of $J^{(N)}_\phi$, namely:
\begin{equation*}
\begin{split}
\lim_{N\to\infty}J^{(N)}_\phi
&=\lim_{N\to\infty}\frac1N\sum_{i=1}^NJ^i(\balpha^1,\cdots,\balpha^N)\\
&=\lim_{N\to\infty}\frac1N\sum_{i=1}^N\EE\bigg[\int_0^Tf(t,X^i_t,\mu^{N}_{t},\phi(t,X^i_t),\nu^{N}_{t})\,dt +g(X^i_T,\mu^{N}_{T})
\bigg],\\
&=\lim_{N\to\infty}\EE\bigg[\int_0^T<f(t,\,\cdot\,,\mu^{N}_{X_t},\phi(t,\,\cdot\,),\nu^{N}_{t}),\mu^N_{t}>\,dt +<g(\,\cdot\,,\mu^{N}_{X_T}),\mu^{N}_{T}>
\bigg],
\end{split}
\end{equation*}
if we use the notation $<\varphi,\rho>$ for the integral $\int \varphi(z)\rho(dz)$ of the function $\varphi$ with respect to the measure $\rho$. Now if we assume that in the limit $N\to\infty$ the empirical distributions $\mu^N_{t}$ converge toward a measure $\mu_t$, and thus $\nu^N_{t}=\frac1N\sum_{i=1}^N\delta_{\phi(t,X^i_t)}$ also converges toward a measure $\nu_t$, then the social cost of the feedback function $\phi$ becomes:
\begin{equation*}
SC(\phi)=\int_0^T<f(t,\,\cdot\,,\mu_t,\phi(t,\,\cdot\,),\nu_t),\mu_t>\,dt \;+\; <g(\,\cdot\,,\mu_T),\mu_T>,
\end{equation*}
with the expectation, $\EE$, disappearing when the limiting flows $\bmu=(\mu_t)_{0\le t\le T}$ and $\bnu=(\nu_t)_{0\le t\le T}$ are deterministic.

@@

\subsection{\textbf{Linear Quadratic Extended Mean Field Games}}\label{sec:EMFG}
To solve the linear quadratic extended mean field game (LQEMFG), we begin by considering the reduced Hamiltonian for this problem:
\begin{equation*}
\begin{split}
    H(t,x,\bar{\mu},\alpha,\bar{\nu},y)=&\left[b_1(t)x+\bar{b}_1(t)\bar{\mu}+b_2(t) \alpha+\bar{b}_2(t)\bar{\nu}\right]y \\
    &+\frac{1}{2}\left[q(t)x^2+\bar{q}(t)(x-s(t)\bar{\mu})^2 +r(t)\alpha^2+\bar{r}(t)(\alpha-\bar{s}(t)\bar{\nu})^2\right],
\end{split}
\end{equation*}
and whenever the flows $\bmubar=(\bar{\mu}_t)_{0\leq t\leq T}$ and $\bnubar=(\bar{\nu}_t)_{0\leq t\leq T}$ are fixed, we consider for each control process $\balpha=(\alpha_t)_{0\leq t\leq T}$ the adjoint equation:
\begin{equation*}
    dY_t=- \partial_x H(t,X_t,\bar{\mu}_t,\alpha_t,\bar{\nu}_t,Y_t)dt+Z_tdW_t,
    \qquad
    Y_T=\partial_xg(X_T, \cL(X_T)).
\end{equation*}
According to the Pontryagin stochastic maximum principle, a sufficient condition for optimality is $\partial_{\alpha}H(t,X_t,\bar{\mu}_t,\hat{\alpha}_t,\bar{\nu}_t,Y_t)=0$. Thus, we find the optimal control:
\begin{equation}
    \hat{\alpha}_t=\frac{\bar{r}(t)\bar{s}(t)\bar{\nu}-b_2(t)Y_t}{r(t)+\bar{r}(t)}.
\label{eq:MFG_opt_alpha_Pontryagin}
\end{equation}
When solving the fixed point step, we identify $\bar{\nu}_t=\mathbb{E}(\hat{\alpha}_t)$. By taking the expectation, we find:
\begin{equation*}
    \bar{\nu}_t = \mathbb{E}(\hat{\alpha}_t)=c^{MFG}(t)\mathbb{E}(Y_t),
	\qquad 
	\text{with} 
	\quad
    c^{MFG}(t)=-\frac{b_2(t)}{r(t)+\bar{r}(t)(1-\bar{s}(t))}.
\end{equation*}
Thus, from equation (\ref{eq:MFG_opt_alpha_Pontryagin}) we have:
\begin{equation}
    \hat{\alpha}_t=a^{MFG}(t) Y_t+b^{MFG}(t)\mathbb{E}(Y_t),
\label{eq:MFG_opt_alpha_Y_expY}
\end{equation}
with:
\begin{equation*}
	a^{MFG}(t)=-\frac{b_2(t)}{r(t)+\bar{r}(t)},
	\qquad
	\text{and}
	\qquad
	b^{MFG}(t)=-\frac{\bar{r}(t)\bar{s}(t)b_2(t)}{(r(t)+\bar{r}(t))(r(t)+\bar{r}(t)(1-\bar{s}(t)))}.
\end{equation*}
Note that $c^{MFG}(t)=a^{MFG}(t)+b^{MFG}(t)$. The solution of the mean field game equilibrium problem is given by the solution to the FBSDE system:
\begin{equation}
\left\{
\begin{array}{lcl}
      \displaystyle  dX_t&=& \left[b_1(t)X_t+\bar{b}_1(t) \mathbb{E}X_t+a^{MFG}(t)b_2(t)Y_t+(b^{MFG}(t)b_2(t)+c^{MFG}(t)\bar{b}_2(t))\mathbb{E}Y_t\right]dt+\sigma dW_t \\[5pt]
        dY_t&=& -\left[(q(t)+\bar{q}(t))X_t-\bar{q}(t)s(t)\mathbb{E}X_t+b_1(t)Y_t \right]dt +Z_t dW_t,
\end{array}
\right.
\label{eq:FBSDE_EMFG}
\end{equation}
with initial condition $X_0=\xi$, a random variable with finite mean and variance, and terminal condition $Y_T=(q_T+\bar{q}_T)X_T-\bar{q}_Ts_T\mathbb{E}X_T$.

This is a linear FBSDE of McKean-Vlasov type, which can be solved explicitly under mild assumptions (or at least in the case of time-independent coefficients which we will consider later. See Appendix \ref{appendix_solving_fbsdes}). Let $\bar{\eta}_t^{MFG}$, $\eta_t^{MFG}$, $\bar{x}_t^{MFG}$, and $v^{MFG}_t$ denote the solutions for this problem as described in the appendix so that:
\begin{align*}
Y_t =\eta_t^{MFG}X_t+(\bar{\eta}_t^{MFG}-\eta_t^{MFG})\bar{x}_t^{MFG}, \quad &\mathbb{E}(Y_t) =\bar{\eta}_t^{MFG}\bar{x}_t^{MFG},\\
\mathbb{E}(X_t) =\bar{x}_t^{MFG}, \quad &Var(X_t)=v^{MFG},
\end{align*}
provide a solution to the LQEMFG problem. Then from the appendix, we have:
\begin{itemize}
	\item a scalar Riccati equation for $\bar{\eta}_t^{MFG}$:
	\begin{equation}
	\dot{\bar{\eta}}^{MFG}_t+ \left[ c^{MFG}(t)(b_2(t)+\bar{b}_2(t)) \right] \cdot (\bar{\eta}^{MFG}_t)^2+ \left(2b_1(t)+\bar{b}_1(t) \right) \cdot \bar{\eta}^{MFG}_t +q(t)+\bar{q}(t)(1-s(t)) =0,
	\label{eq:eta_bar_MFG}
	\end{equation}
	with terminal condition $\bar{\eta}^{MFG}_T = q_T+\bar{q}_T(1-s_T)$, 
	\item a linear first order ODE for $\bar{x}_t^{MFG}$:
	\begin{equation}
	\dot{\bar{x}}^{MFG}_t=\left[ b_1(t)+\bar{b}_1(t)+c^{MFG}(t)(b_2(t)+\bar{b}_2(t)) \cdot \bar{\eta}^{MFG}_t \right] \cdot \bar{x}^{MFG}_t,
	\label{eq:x_bar_MFG}
	\end{equation}
	with initial condition $\bar{x}^{MFG}_0=\mathbb{E}(\xi)$,
	\item a scalar Riccati equation for $\eta_t^{MFG}$:
	\begin{equation}
	\dot{\eta}^{MFG}_t+a^{MFG}(t)b_2(t)\cdot (\eta^{MFG}_t)^2 + 2b_1(t) \cdot \eta^{MFG}_t+q(t)+\bar{q}(t)=0,
	\end{equation}
	with terminal condition $\eta^{MFG}_T = q_T+\bar{q}_T$,
\end{itemize}
and where the dot is the standard ODE notation for a derivative. 
And thus, we obtain explicit solutions for $\bar{x}_t^{MFG}$ and $v_t^{MFG}$:
\begin{IEEEeqnarray}{rCl}
	\bar{x}^{MFG}_t &=& \mathbb{E}(\xi) e^{ \int_0^t\left( b_1(s)+\bar{b}_1(s)+ \left[ c^{MFG}(s)(b_2(s)+\bar{b}_2(s))\right] \cdot \bar{\eta}^{MFG}_s\right) ds} \label{eq:x_bar_MFG_explicit}, \\ 
	v^{MFG}_t &=& Var(\xi)e^{\int_0^t 2\left[b_1(s)+a^{MFG}(s)b_2(s) \cdot \eta^{MFG}_s\right]ds}+\sigma^2 \int_0^t e^{2 \int_s^t \left[b_1(u)+a^{MFG}(u)b_2(u)\eta^{MFG}_u\right] du}ds.
	\label{eq:v_t}
\end{IEEEeqnarray}
Let $SC^{MFG}:=SC(\phi)$ in which $\phi=\phi(t,x)$ is the feedback function specified by this solution, namely, from equation (\ref{eq:MFG_opt_alpha_Y_expY}) we have:
\begin{equation*}
    \phi(t,x)=a^{MFG}(t)\eta_t^{MFG}x+\left[a^{MFG}(t)(\bar{\eta}_t^{MFG}-\eta_t^{MFG})+b^{MFG}(t)\bar{\eta}_t^{MFG} \right]\bar{x}_t^{MFG}.
\end{equation*}
Then we can compute the social cost as described in Section \ref{sub:MFG_formulation}:
\begin{equation*}
\begin{split}
    SC^{MFG}=&\frac{1}{2}\Big[(q_T+\bar{q}_T)v_T^{MFG}+(q_T+\bar{q}_T(1-s_T)^2)(\bar{x}_T^{MFG})^2\\
    &+\int_0^T \left[q(t)+\bar{q}(t)+(r(t)+\bar{r}(t))(a^{MFG}(t)\eta_t^{MFG})^2\right]v_t^{MFG}dt\\
    &+\int_0^T\left[q(t)+\bar{q}(t)(1-s(t))^2+(r(t)+\bar{r}(t)(1-\bar{s}(t))^2)(c^{MFG}(t)\bar{\eta}_t^{MFG})^2\right](\bar{x}_t^{MFG})^2dt\Big],
\end{split}
\end{equation*}
where we have used the fact that:
\begin{equation*}
    \mathbb{E}(\phi(t,X_t))=c^{MFG}(t)\cdot \bar{\eta}_t^{MFG}\bar{x}_t^{MFG}
	\qquad 
	\text{and}
	\qquad
    Var(\phi(t,X_t))=\left(a^{MFG}(t)\eta_t^{MFG} \right)^2 \cdot v^{MFG}_t.
\end{equation*}

\subsection{\textbf{Linear Quadratic Control of McKean-Vlasov Type Involving the Law of the Control}}\label{sec:EMKV}
To solve the linear quadratic optimal control problem of McKean-Vlasov type involving the law of the control (LQEMKV), we begin with the reduced Hamiltonian, which is the same as in the LQEMFG problem:
\begin{equation*}
\begin{split}
    H(t,x,\bar{\mu},\alpha,\bar{\nu},y)=&\left[b_1(t)x+\bar{b}_1(t)\bar{\mu}+b_2(t) \alpha+\bar{b}_2(t)\bar{\nu}\right]y\\
    &+\frac{1}{2}\left[q(t)x^2+\bar{q}(t)(x-s(t)\bar{\mu})^2 +r(t)\alpha^2+\bar{r}(t)(\alpha-\bar{s}(t)\bar{\nu})^2\right].
\end{split}
\end{equation*}
Since we require $\bar{\nu}_t$ to be equal to $\mathbb{E}(\alpha_t)$ throughout the optimization, it is not sufficient to minimize the Hamiltonian with respect to the $\alpha$ input alone in order to guarantee optimality. A sufficient condition for control problems of McKean-Vlasov type involving the law of the control is derived in \cite{carmona_acciaio}. Since we consider a Hamiltonian that depends on the means of $\bar{\mu}$ and $\bar{\nu}$ instead of the full distributions, the sufficient condition reduces to the following (see section 4 in \cite{carmona_acciaio}):
\begin{equation*}
    \partial_{\alpha}H(t,X_t,\mathbb{E}(X_t),\hat{\alpha}_t,\mathbb{E}(\hat{\alpha}_t),Y_t)+\tilde{\mathbb{E}} \left[\partial_{\bar{\nu}}H(t,\tilde{X}_t,\mathbb{E}(X_t),\hat{\alpha}_t,\mathbb{E}(\hat{\alpha}_t),\tilde{Y}_t) \right]=0,
\end{equation*}
where the adjoint equation is given by:
\begin{equation*}
\left\{
\begin{array}{lcl}
    dY_t&=&-\left[\partial_x H(t,X_t,\bar{\mu}_t,\alpha_t,\bar{\nu}_t,Y_t)+\tilde{\mathbb{E}} \left[\partial_{\bar{\mu}}H(t,\tilde{X}_t,\bar{\mu}_t,\tilde{\alpha}_t,\bar{\nu}_t,\tilde{Y}_t) \right] \right]dt+Z_tdW_t \\[8pt]
    Y_T&=&\partial_xg(X_T, \cL(X_T))+\tilde{\mathbb{E}} \left[\partial_{\bar{\mu}}g(\tilde{X}_T,\cL(X_T))(X_T) \right],
\end{array}
\right.
\end{equation*}
and where $(\boldsymbol{\tilde{X}},\boldsymbol{\tilde{Y}},\boldsymbol{\tilde{\alpha}})$ denotes an independent copy of $(\boldsymbol{X},\boldsymbol{Y},\boldsymbol{\alpha})$. In the present LQ case, the sufficient condition can be used to solve for:
\begin{equation}
    \hat{\alpha}_t=a^{MKV}(t)Y_t+b^{MKV}(t)\mathbb{E}(Y_t),
    \label{eq:alpha_hat_MKV}
\end{equation}
with:
\begin{equation*}
\begin{split}
    a^{MKV}(t)&=-\frac{b_2(t)}{r(t)+\bar{r}(t)}, \quad \text{and} \quad b^{MKV}(t)=-\frac{1}{r(t)+\bar{r}(t)}\left(\bar{b}_2(t)-\frac{\bar{r}(t)\bar{s}(t)(\bar{s}(t)-2)(b_2(t)+\bar{b}_2(t))}{r(t)+\bar{r}(t)(1-\bar{s}(t))^2} \right).
\end{split}
\end{equation*}
Then $\mathbb{E}(\hat{\alpha}_t)=c^{MKV}(t)\mathbb{E}(Y_t)$ with:
   $$ c^{MKV}(t)=a^{MKV}(t)+b^{MKV}(t)=-\frac{b_2(t)+\bar{b}_2(t)}{r(t)+\bar{r}(t)(1-\bar{s}(t))^2}.$$

@@

\subsection{\textbf{A Particular Example: Flocking}}
Mean field game models of flocking have been proposed in the literature by Nourian et al \cite{nourian2010synthesis}\cite{nourian2011mean}. Here we consider a slightly different formulation as described in Section 3.6.1 of the book of Carmona and Delarue \cite{Carmona_book}. A representative bird in the flock controls their velocity, $X_t$, through the drift:
\begin{equation*}
    b(t,x,\mu,\alpha,\nu)=\alpha.
\end{equation*}
Suppose the initial condition is $X_0=1$ and the volatility is $\sigma=1$. They choose their control with two goals in mind: to minimize their effort put into their control, and to align their velocity with the average velocity of the group. Thus, they consider the cost functions given by:
\begin{equation*}
\begin{split}
    f(t,x,\mu,\alpha,\nu)=\frac{1}{2}\left(|x-\bar{\mu}|^2 +\alpha^2 \right), \quad \text{and} \quad g(x,\mu)=0.
\end{split}
\end{equation*}
